Viscous Erosion in Porous Media

### 2018-03-16 docs/methods/erosion_methods.tex
NM

	The drag exerted by the eroding body offers a second check on the simulation. Figure~\ref{area_drag}(b) shows the horizontal drag, $F_D$, calculated in the simulations (black curve) using Eq.~\eqref{eqn:drag}. In the context of slender-body theory, $F_D$ should be interpreted as drag-per-unit-span. As seen in Fig.~\ref{area_drag}(b), the drag decreases gradually at first and then abruptly tends to zero in the very final moments. This behavior, though perhaps surprising, is consistent with the fact that drag vanishes {\em logarithmically} with body-size in 2D Stokes flow. Equation \eqref{eqn:drag} also allows us to calculate the separate contributions of pressure and viscous drag, and we show these with the blue and green curves in Fig.~\ref{area_drag}(b). Initially, pressure and viscous drag are comparable, with the circular geometry having a slightly higher level of pressure drag. However,  as the body erodes and assumes a more slender form, the pressure drag decreases much more rapidly, leaving viscous drag to dominate. This behavior is associated with the tendency towards a drag-minimizing profile, which will be discussed in Section \ref{LimitingShape}.

As a check on these results, we can use slender-body theory to estimate the drag-per-unit-span
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}	\qquad \text{for } a \ll \ell \, .
\end{align}
For the case of a {\em circular} cylinder, the normalization factor of $4 \pi$ produces an {\em exact} relationship (under the slender-body assumptions), where $a$ is taken to be the radius. For other shapes the formula will be approximate and we define $a = \sqrt{A/\pi}$ as the characteristic lengthscale (since it generalizes the case of a circle). In Fig.~\ref{area_drag}(b) we show this scaling prediction (red dotted curve), where $\ell = 2.3 a_0$ is a fit parameter and $a_0 = 0.2$ the radius of the initially circular body. The scaling law agrees closely with the simulated drag as the body vanishes, thus providing validation.

>>>>

	The drag exerted by the eroding body offers a second check on the simulation. Figure~\ref{area_drag}(b) shows the horizontal drag, $F_D$, calculated in the simulations (black curve) using Eq.~\eqref{eqn:drag}. In the context of slender-body theory, $F_D$ should be interpreted as drag-per-unit-span. As seen in Fig.~\ref{area_drag}(b), the drag decreases gradually at first and then abruptly tends to zero in the very final moments. This behavior, though perhaps surprising, is consistent with the fact that drag vanishes {\em logarithmically} with body-size in 2D Stokes flow. Equation \eqref{eqn:drag} also allows us to calculate the separate contributions of pressure and viscous drag, and we show these with the blue and green curves in Fig.~\ref{area_drag}(b). Initially, pressure and viscous drag are comparable, with the circular geometry having a slightly higher level of pressure drag. As the body erodes and assumes a more slender form, though, the pressure drag decreases much more rapidly, leaving viscous drag to dominate. This behavior is associated with the tendency towards a drag-minimizing profile, which will be discussed in Section \ref{LimitingShape}.

As a check on these results, we can use slender-body theory to estimate the drag-per-unit-span
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}	\qquad \text{for } a \ll \ell \, .
\end{align}
For the case of a {\em circular} cylinder, the normalization factor of $4 \pi$ produces an {\em exact} relationship (under the slender-body assumptions), with $a$ taken to be the radius. For other shapes the formula will be approximate and we define $a = \sqrt{A/\pi}$ as the characteristic lengthscale (since it generalizes the case of a circle). In Fig.~\ref{area_drag}(b) we show this scaling prediction (red dotted curve), where $\ell = 2.3 a_0$ is a fit parameter and $a_0 = 0.2$ is the radius of the initially circular body. The scaling law and measured drag agree closely as the body vanishes.
